---
title: "Literature review"
author: "Juho Salminen"
date: "4 May 2015"
output: html_document
---


```{r,echo=FALSE, message=FALSE, results='hide'}
# Loading libraries
library(ggplot2)
library(splitstackshape)
library(igraph)
library(knitr)

# Set ggplot theme
theme_set(theme_minimal(15))
options(stringsAsFactors = FALSE)
```


```{r, echo=FALSE, warning=FALSE, results='hide'}
# Loading and preparing data

# Call cleaning2.R to process the data in the input folder and 
# save processed files to output folder
source("scopus_cleaning.R", chdir = T)

# Load data
literature <- read.csv("output/literature.csv", header = T, 
                       sep = ";")
literatureByAuthor <- read.csv("output/literature_by_author.csv", header = T, 
                               sep = ";", )
literatureByKeywords <- read.csv("output/literature_by_keywords.csv", header = T, 
                               sep = ";")
literatureByIndex <- read.csv("output/literature_by_index.csv", header = T, 
                               sep = ";")
citationEdges <- read.csv("output/citation_edges.csv", header = T, 
                          sep = ";")
citationNodes <- read.csv("output/citation_nodes.csv", header = T, 
                          sep = ";")

# Helper function to remove leading and trailing whitespace
trim <- function (x) gsub("^\\s+|\\s+$", "", x)

# Fixing variable types
```

## Data description
Number of dimensions (rows and columns) and variable names. 
```{r, echo=FALSE}
# Describing the data
dims <- dim(literature)
cat("Data consist of", dims[1], "observations of", dims[2], "variables.")
```

## Distribution of publication years 
```{r, echo=FALSE}
qplot(data = literature, Year, fill = I("darkgreen"), binwidht = 1) + 
    ggtitle("Year published") + 
    xlab("Year") + 
    ylab("Article count")
summary(literature$Year)
```

## Distribution of citation counts
The lowest 99 % of citation counts are shown. There are often outliers with 
orders of more citations than the typical articles.
```{r,echo=FALSE}
# Find 99th quantile of times cited and create plot title
q99 <- quantile(literature$Cited.by, .99, na.rm = T)
titleText <- paste("Times cited\n(< ", floor(q99), " citations)", sep = "")

qplot(data = literature, Cited.by, binwidth = 1, fill = I("red")) + 
    scale_x_continuous(limits = c(0,q99)) + 
    ggtitle(titleText) + 
    xlab("Number of times cited") + 
    ylab("Article count")

summary(literature$Cited.by)
```

## Important authors
```{r,echo=FALSE, fig.height=7}
# Calculating total number of citations for each author
citationSums <- aggregate(literatureByAuthor$Cited.by, 
    by = list(Authors = toupper(literatureByAuthor$Authors)), 
    FUN = sum)

# Fixing column names
names(citationSums) <- c("Authors", "Cited.by")
# Crating new data frame to plot citations by author

# Extract author names 
authorNames <- unlist(strsplit(literature$Authors, ","))
# Remove apostrophes
authorNames <- gsub("'", "", authorNames)
# Count author name frequencies
authors <- table(authorNames)
# Transform to a data frame
authors <- as.data.frame(authors)
# Merge with data frame containing the total times citated by each author
authors <- merge(authors, citationSums, by.x = "authorNames", 
              by.y = "Authors" )
# Fix column name
names(authors)[1] <- "Authors" 
# Sort the table by total times sited, decreasing order
authors <- authors[with (authors, order(-Cited.by)), ]
# Extract the names of top 25 most cited authors
top25 <- head(authors$Authors, 25)
# Merge with the rest of the data. This data frame can have the same author name
# multiple times
authorsdf <- merge(authors, literatureByAuthor, by = "Authors")
# Extract top 25 most cited authors. This data frame has each author name only once.
top25Authors <- head(authors, 25)
# Top 25 authors again. Each article on a separate row.
top25Authorsdf <- authorsdf[authorsdf$Authors %in% top25, ]
top25Authorsdf <- transform(top25Authorsdf, 
                            Authors = reorder(Authors, Cited.by.y))

# Sort authors by number of articles, extract top 25, 
# and reorder factors for plotting
authors <- authors[with (authors, order(-Freq)), ]
authorsPop <- head(authors, 25)
authorsPop <- transform(authorsPop, Authors = reorder(Authors, Freq))

ggplot(authorsPop, aes(Authors, Freq)) +
    geom_bar(stat = "identity", fill = "blue") + 
    coord_flip() +
    ggtitle("Top 25 most productive authors") + 
    xlab("Author") + 
    ylab("Number of articles")
```

Sorted by number of articles published.

```{r,echo=FALSE}
ggplot(top25Authorsdf, aes(Authors, Cited.by.y)) + 
    geom_boxplot(color = "blue") +
    coord_flip() +
    ggtitle("Citations of top 25\nmost cited authors") + 
    xlab("Author") + ylab("Times cited")
```

Sorted by total number of citations.

## Important publications
```{r,echo=FALSE, fig.width=10, fig.height=7}
# Calculating total citations for each publication.
# Functionality same as for the authors, see above.

citationSums <- aggregate(literature$Cited.by, 
    by = list(Source.title= literature$Source.title), 
    FUN = sum)
names(citationSums) <- c("Source.title", "PublicationTotalCitations")
citationSums <- citationSums[with (citationSums, order(-PublicationTotalCitations)), ]
top25 <- head(citationSums[,c("Source.title", "PublicationTotalCitations")], 25)

publications <- sort(table(literature$Source.title), decreasing = T)
publications <- as.data.frame(publications)
publications$Publication <- row.names(publications)
names(publications) <- c("Count", "Publication")
publications$Publication <- strtrim(publications$Publication, 50)
publications <- transform(publications, Publication = reorder(Publication, Count))

literature <- merge(literature, citationSums, 
                    by = "Source.title" )

top25Lit <- literature[literature$Source.title %in% top25$Source.title, ]
top25Lit$Source.title <- strtrim(top25Lit$Source.title, 50)
top25Lit <- transform(top25Lit,  
            Source.title = reorder(Source.title, PublicationTotalCitations))

ggplot(head(publications, 25), aes(Publication, Count)) + 
    geom_bar(stat = "identity", fill = "orange") + 
    coord_flip() + 
    theme(legend.position = "none") +
    ggtitle("Most popular publications") + 
    xlab("Publication") + 
    ylab("Article count")
```

Sorted by number of published articles in the dataset.

```{r,echo=FALSE, fig.width=10, fig.height=7}
ggplot(top25Lit, 
       aes(Source.title, Cited.by)) + 
    geom_boxplot(color = "orange") + 
    coord_flip() +
    theme(legend.position = "none") +
    ggtitle("Top 25\nmost cited publications") + 
    xlab("Publication") + ylab("Times cited")
```

Sorted by the total number of citations.

## Important keywords

```{r, echo=FALSE, fig.height=7, fig.width=8}
# Calculating total citations for each keyword
# Functionality same as for the authors, see above.

keywordCitationSum <- aggregate(literatureByKeywords$Cited.by, 
                                by = list(Author.Keywords = 
                            literatureByKeywords$Author.Keywords), FUN = sum)
names(keywordCitationSum) <- c("Author.Keywords", "TotalTimesCited")

keywords <- unlist(strsplit(literature$Author.Keywords, ";"))
keywords <- table(keywords)
keywords <- as.data.frame(keywords)
names(keywords) <- c("Author.Keywords", "Freq")

keywords <- merge(keywords, keywordCitationSum, by = "Author.Keywords")

keywords <- keywords[with (keywords, order(-TotalTimesCited)), ]
top25 <- head(keywords$Author.Keywords, 25)
keywordsdf <- merge(keywords, literatureByKeywords, by = "Author.Keywords")
top25keywords <- head(keywords, 25)
top25keywordsdf <- keywordsdf[keywordsdf$Author.Keywords %in% top25, ]

keywords <- keywords[with (keywords, order(-Freq)), ]
keywordsPop <- head(keywords, 25)
keywordsPop <- transform(keywordsPop, Author.Keywords = 
                             reorder(Author.Keywords, Freq))

top25keywordsdf <- transform(top25keywordsdf, 
                             Author.Keywords = reorder(Author.Keywords, TotalTimesCited))

ggplot(keywordsPop, aes(Author.Keywords, Freq)) +
    geom_bar(stat = "identity", fill = "purple") + 
    coord_flip() +
    ggtitle("Top 25 most popular keywords") + 
    xlab("Keyword") + 
    ylab("Number of occurences")
```

Sorted by number of articles where the keyword is mentioned.

```{r,echo=FALSE}
ggplot(top25keywordsdf, aes(Author.Keywords, Cited.by)) + 
    geom_boxplot(color = "purple") +
    coord_flip()  +
    ggtitle("Citations of top 25 most cited keywords") + 
    xlab("Keyword") + ylab("Times cited")
```

Sorted by total number of citations for the keyword.

## The most important papers
The most important papers and other sources are identified below using several importance measures.

### The most cited articles
Total citation counts according to Web of Science.
```{r,echo=FALSE}
# Remove edges with nodes missing from node table
correct <- (citationEdges$Target %in% citationNodes$Id & 
                 citationEdges$Source %in% citationNodes$Id)
citationEdges <- citationEdges[correct, ]

# Create igraph
citationGraph <- graph.data.frame(citationEdges, vertices = citationNodes)
# Calculate PageRanks
citationNodes$PageRank <- page.rank(citationGraph)$vector
# Calculate in-degrees
citationNodes$InDegree <- degree(citationGraph, mode = "in")

# Extract the articles included in the data set and articles not included
# in the dataset
citationsLit <- citationNodes[citationNodes$Origin == "literature", ]
citationsLit <- merge(citationsLit, literature[, c(3, 11, 16, 42, 43)], 
                       by.x = "Id", by.y = "ReferenceString")
citationsLit$Article <- paste(toupper(citationsLit$Title), " | ",
                              citationsLit$Id, " | ", 
                                      citationsLit$Abstract)
citationsRef <- citationNodes[citationNodes$Origin == "reference", ]

# Order by TimesCited, decreasing order
citationsLit <- citationsLit[with (citationsLit, order(-Cited.by)), ]
kable(head(citationsLit[, c("Article", "Cited.by")], 25))
```

### The papers with the highest in-degree
Literature in the dataset references these sources the most often. The following papers are included in the dataset.
```{r,echo=FALSE}
# Order by InDegree, decreasing order
citationsLit <- citationsLit[with (citationsLit, order(-InDegree)), ]
citationsRef <- citationsRef[with (citationsRef, order(-InDegree)), ]
# Display the top 25 of most cited references included in the dataset 
kable(head(citationsLit[, c("Article", "InDegree")], 25))
```


### The references with the highest in-degree NOT found in the dataset
Literature in the dataset references these sources the most often. These papers are not included in the dataset and might warrant inclusion.
```{r,echo=FALSE}
# Display the top 25 of most cited references not included in the dataset 
kable(head(citationsRef[, c("Id", "InDegree")], 25))
```

### The articles with highest PageRank scores
PageRank score calculated within the citation network of the dateset.
```{r,echo=FALSE}
# Order by PageRank decreasing order
citationsLit <- citationsLit[with (citationsLit, order(-PageRank)), ]
citationsRef <- citationsRef[with (citationsRef, order(-PageRank)), ]

# Display the top 25 papers with highest PageRank included in the dataset 
kable(head(citationsLit[, c("Article", "PageRank")], 25))
```

### The references with the highest PageRank scores NOT found in the dataset
PageRank score calculated within the citation network of the dateset.
```{r,echo=FALSE}
# Display the top 25 papers with highest PageRank not included in the dataset 
kable(head(citationsRef[, c("Id", "PageRank")], 25))
```


